### 2024-12-26 skripsi/chapter4.tex
finish ch4 scenaro 1 results

\begin{figure}[H]
  \centering
  \begin{adjustbox}{max width=\textwidth}
    \tikzfig{figures/computational_model}
  \end{adjustbox}
  \caption{Computational Model of SpectralSVR}\label{fig:computational_model}
\end{figure}

@@

[build-system]
requires = ["poetry-core"]
build-backend = "poetry.core.masonry.api"
  \end{lstlisting}
  \caption{Contents of pyproject.toml configuration file that define the dependencies using Poetry}\label{fig:poetry_add_deps}
\end{figure}

@@

def to_complex_coeff(x: torch.Tensor) -> torch.Tensor:
  return torch.view_as_complex(
    x.reshape((*x.shape[:-1], -1, 2))
  )
  \end{lstlisting}
  \caption{Utility function to convert between complex matrices and the real representations}\label{fig:complex_matrix_conversion_implementation}
\end{figure}

@@

    x_resized = torch.concat(
        (
            x_resized.index_select(dim, start_range),
            x_resized.index_select(dim, end_range),
        ),
        dim,
    )
  \end{lstlisting}
  \caption{Example of shrinking a tensor of samples' coefficients to a target number of modes}\label{fig:coefficient_shrink_impl}
\end{figure}
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
for dim, (target_mode, current_mode) in enumerate(
    zip(target_modes, current_modes), 1
):
    device = x.device
    start_range = torch.tensor(
        range((current_mode - 1) // 2 + 1), dtype=torch.int
    ).to(device=device)
    # make sure that end range is empty if
    # the coefficient is only size 1
    end_range = torch.tensor(
        range(current_mode // 2, current_mode)
        if current_mode > 1
        else range(0),
        dtype=torch.int,
    ).to(device=device)
    padding_size = target_mode - current_mode
    modes = list(x_resized.shape)
    modes[dim] = padding_size
    padding = torch.zeros(modes).to(x_resized)

    x_resized = torch.concat(
        (
            x_resized.index_select(dim, start_range),
            padding,
            x_resized.index_select(dim, end_range),
        ),
        dim,
    )
  \end{lstlisting}
  \caption{Example of expanding a tensor of samples' coefficients to a target number of modes}\label{fig:coefficient_expand_impl}
\end{figure}
Next, the second way to manage the size of a tensor along a dimension is to interpolate the current values. The implementation uses simple linear interpolation. This is especially useful for situations such as coefficients that are themselves functions of time. This is a common application such as when solving the initial condition problem of a function in space and time. One might represent the function as a Fourier series with basis as functions of space and coefficients as functions of time. In this case, the coefficients for time values that are not available can be interpolated. The implementation of this is shown in \lccref{fig:coefficient_interpolation_impl}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
    index_floor = index_float.floor().to(torch.int)
    index_ceil = index_float.ceil().to(torch.int)
    x_ceil = x.index_select(dim, index_ceil)
    x_floor = x.index_select(dim, index_floor)
    # interpolate coefficients
    index_shape = [1 for _ in range(x_floor.ndim)]
    index_shape[1] = -1
    index_scaler = (
        ((index_float - index_floor) / (
          index_ceil - index_floor
        ))
        .reshape(index_shape)
        .nan_to_num()
    )
    # ynt + scaler * (ynt1 - ynt)
    # (1 - scaler) * ynt + scaler * ynt1
    x_interp = torch.lerp(x_floor, x_ceil, index_scaler.to(x))
  \end{lstlisting}
  \caption{Example of interpolating for indices (floating point indices) between the available ones (intiger indices)}\label{fig:coefficient_interpolation_impl}
\end{figure}

@@

lsoda_solver: SolverSignatureType = partial(
    odeint, method="scipy_solver", options={"solver": "LSODA"}
)  # type: ignore
  \end{lstlisting}
  \caption{Implementation of ODE solvers and the types}\label{fig:solver_impl}
\end{figure}

@@

        return F
...
  \end{lstlisting}
  \caption{Implementation of forward Fourier transform}\label{fig:fourier_transform_impl}
\end{figure}

@@

        return F
...
  \end{lstlisting}
  \caption{Implementation of n-dimensional Fourier transform for a specific dimension}\label{fig:ndim_fourier_transform_impl}
\end{figure}

@@

            F = torch.mm(f, e.T)
        return F
...
  \end{lstlisting}
  \caption{Implementation of the one dimensional Fourier Transform}\label{fig:1d_fourier_transform_impl}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def __add__(self, other: Self):
        if isinstance(other, self.__class__):
            if other.coeff is None:
                return self.copy()
            elif self.coeff is None:
                return other.copy()
            else:
                result = self.resize_modes(other)
                result.coeff = result.coeff + other.coeff
                return result
        else:
            raise TypeError(
                f"unsupported operand type(s) for +: '{
                    self.__class__
                }' and '{type(other)}'"
            )
...
  \end{lstlisting}
  \caption{Implementation of basis addition function}\label{fig:basis_addition_impl}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def grad(self, dim: int = 0, ord: int = 1) -> Self:
        copy = self.copy()
        if dim == 0 and self.time_dependent:
            # time dependent use finite differences
            dt = self.periods[0] / (self.time_size - 1)
            coeff = copy.coeff
            for o in range(ord):
                coeff = torch.gradient(coeff, spacing=dt, dim=1)[0]
            copy.coeff = coeff
        else:
            if self.time_dependent:
                # disregard time dimension
                dim = dim - 1
            k = copy.wave_number(copy.modes[dim])
            multiplier_dims = [1 for _ in range(copy.ndim)]
            multiplier_dims[dim] = copy.modes[dim]
            if self.time_dependent:
                multiplier_dims = (1, *multiplier_dims)
            multiplier = (
                2
                * torch.pi
                * 1j
                * k.reshape(multiplier_dims).to(copy.coeff)
                / self.periods[dim]
            )
            multiplier = multiplier.pow(ord)
            coeff = copy.coeff.mul(multiplier)
            coeff[:, ..., 0] = torch.tensor(0 + 0j)
            copy.coeff = coeff
        return copy
...
  \end{lstlisting}
  \caption{Implementation of basis grad function}\label{fig:basis_grad_impl}
\end{figure}

@@

            vals = cls.inv_transform(coeff)
            coeff = cls.transform(vals.real + 0j)
        if scale:
            scaler = torch.tensor(modes).sum() * 0.2
            coeff = coeff.mul(scaler)
        return coeff
...
  \end{lstlisting}
  \caption{Implementation of Fourier Basis coefficient generation function}\label{fig:generate_coeff_impl}
\end{figure}

@@

\subsubsection{Model}

@@

        return (b, alpha)
...
  \end{lstlisting}
  \caption{Implementation of Least-squares Support Vector Regression fitting function}\label{fig:lssvr_fit_function_impl}
\end{figure}

@@

        if X.ndim == 1:
            return predictions.reshape(-1)
        else:
            return predictions
...
  \end{lstlisting}
  \caption{Implementation of Least-squares Support Vector Regression prediction function}\label{fig:lssvr_predict_function_impl}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def get_correlation_image(self):
        return self._batched_K(
            self.sv_x, self.sv_x
        ).mm(self.sv_x)
...
  \end{lstlisting}
  \caption{Implementation of Least-squares Support Vector Regression correlation image function}\label{fig:lssvr_correlation_image_function_impl}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def get_p_matrix(self):
        return self.sv_x.T.mm(self.alpha)
...
  \end{lstlisting}
  \caption{Implementation of Least-squares Support Vector Regression p-matrix function}\label{fig:lssvr_p_matrix_function_impl}
\end{figure}

@@

        if torch.is_complex(u_coeff):
            u_coeff = to_real_coeff(u_coeff)
        if torch.is_complex(f):
            f = to_real_coeff(f)
        self.svr.fit(f, u_coeff)
        return self
...
  \end{lstlisting}
  \caption{Implementation of SpectralSVR training function}\label{fig:spectralsvr_train_impl}
\end{figure}

@@

        return self.basis.evaluate(
            coeff=coeff.reshape((f.shape[0], *self.modes)),
            x=x,
            periods=periods,
            time_dependent=self.basis.time_dependent,
        )
...
  \end{lstlisting}
  \caption{Implementation of SpectralSVR forward function (pointwise prediction)}\label{fig:spectralsvr_forward_impl}
\end{figure}

The final core piece of functionality the SpectralSVR class provides is the inverse parameter estimation. This is implemented to solve inverse problems such as predicting the initial conditions for the Burgers' equation or heat equation. This function essentially does the gradient descent on a loss function in such a way that the output function matches some expected output. The loss function can simply be the coefficients of the expected output function. The loss function is formulated as the mean squared error of the difference between the predicted output and expected output. Using some randomly initialized values as the inputs, the predicted output is then used to compute the loss. The gradient of the loss with respect to the inputs are then computed and then used to perform optimization of the inputs using the ADAM optimizer. This is implemented as the inverse function shown in \lccref{fig:spectralsvr_inverse_impl}. The function arguments are the expected output function coefficients, evaluation points, the loss function used, how many optimization loops are done, the random number generator, and the gain which is used in generating the random initial input functions. The function returns the estimated input function coefficients in the real representation. For Fourier coefficients this means that the predicted input coefficients need to first be converted to complex valued tensors and then used to construct a FourierBasis instace.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def inverse(
        self,
        u: torch.Tensor,
        points: torch.Tensor,
        loss_fn: Callable[
            [torch.Tensor, torch.Tensor], torch.Tensor
        ] = mean_squared_error,
        epochs=100,
        generator=torch.Generator().manual_seed(42),
        gain=0.2,
        **optimizer_params,
    ):
        f_shape = (u.shape[0], self.svr.sv_x.shape[1])
        # inverse problem
        f_pred = torch.randn(f_shape, generator=generator) * gain
        f_pred.requires_grad_()

        optim = torch.optim.Adam([f_pred], **optimizer_params)
        for epoch in range(epochs):
            optim.zero_grad()
            u_pred = self.forward(f_pred, points).real
            loss = loss_fn(u_pred, u)
            loss.backward()
            optim.step()
        optim.zero_grad()
        return f_pred
...
  \end{lstlisting}
  \caption{Implementation of SpectralSVR inverse function (parameter estimation)}\label{fig:spectralsvr_inverse_impl}
\end{figure}

The last piece of the SpectralSVR class provides a convenience function for testing the performance of the learned model. This function is shown in \lccref{fig:spectralsvr_test_impl}. The function arguments are the testing inputs, testing outputs, and a resolution for evaluation of function values. The first step is to convert the testing input into real representations and then predicting the output function. The predicted output are then compared to the testing outputs using many metrics including RMSE, MSE, MAE, R2, SMAPE, RSE, RRSE, and the number of Nan values present in the predicted output. The same comparison process is carried out again for the function values which are evaluated at the specified resolution. Finally, the function returns the metric values as a nested tuple for both the coefficients predictions and the function value predictions.

@@

        grid = self.basis.grid(res).flatten(0, -2)
        u_preds = self.basis.evaluate(
            coeff=to_complex_coeff(u_coeff_preds),
            x=grid,
            time_dependent=self.basis.time_dependent,
        ).real
        u_targets = self.basis.evaluate(
            coeff=to_complex_coeff(u_coeff_targets),
            x=grid,
            time_dependent=self.basis.time_dependent,
        ).real
        return {
            "spectral": get_metrics(
                u_coeff_preds, u_coeff_targets
            ),
            "function value": get_metrics(u_preds, u_targets),
        }
...
  \end{lstlisting}
  \caption{Implementation of SpectralSVR test function}\label{fig:spectralsvr_test_impl}
\end{figure}

@@

    @abc.abstractmethod
    def residual(
        self, u: BasisSubType, *args, **kwargs
    ) -> BasisSubType:
        pass
  \end{lstlisting}
  \caption{Implementation of Problem base class}\label{fig:problem_impl}
\end{figure}

@@

        return (u, ut)
...
  \end{lstlisting}
  \caption{Implementation of Antiderivative generate function}\label{fig:antiderivative_generate_impl}
\end{figure}

@@

    def residual(
        self, u: BasisSubType, ut: BasisSubType
    ) -> BasisSubType:
        u_val, grid = u.get_values_and_grid()
        ut_val = ut.get_values()
        dt = grid[1, 0] - grid[0, 0]
        u_grad = torch.gradient(
            u_val, spacing=dt.item(), dim=1
        )[0]
        residual_val = u_grad - ut_val
        residual = u.copy()
        residual.coeff = u.transform(residual_val)
        return residual
...
  \end{lstlisting}
  \caption{Implementation of Antiderivative residual functions}\label{fig:antiderivative_residual_impl}
\end{figure}

@@

        L = space_domain.stop - space_domain.start
        x = (
            basis.grid(slice(
                space_domain.start, space_domain.stop, modes[0]
            )).flatten(0, -2)
            .to(device=device)
        )
        T: float = time_domain.stop - time_domain.start
        nt = int(time_domain.step)
        # nt = int(T / (0.01 * nu) + 2)
        dt = T / (nt - 1)
        t = basis.grid(time_domain).flatten().to(device=device)
        periods = (T, L)
...
  \end{lstlisting}
  \caption{Implementation of Burgers generate function}\label{fig:burgers_generate_impl}
\end{figure}

@@

                f_val = fst.get_values(
                    res=(time_domain, *res_modes)
                )
                f_coeff = basis.transform(
                    f_val.flatten(0, 1)
                ).reshape((n, nt, *modes))
                f_gen = basis(
                    coeff=f_coeff,
                    time_dependent=True,
                    periods=periods
                )
...
  \end{lstlisting}
  \caption{Implementation of Burgers generate function manufactured method}\label{fig:burgers_generate_manufactured_impl}
\end{figure}

@@

        u_u_x_hat = uu_x.coeff
        u_xx_hat = u.grad().grad().coeff
        u_t_hat = nu * u_xx_hat + f_hat - u_u_x_hat
        return u_t_hat
...
  \end{lstlisting}
  \caption{Implementation of Burgers generate function time derivative}\label{fig:burgers_generate_dt_impl}
\end{figure}
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
        u0.coeff = u0.coeff.to(device=device)
        fst.coeff = fst.coeff.to(device=device)
        print(f"generating with {len(t)} time steps")

@@

        u_hat = u_hat.movedim(0, 1)
        if timedependent_solution:
            u = basis(u_hat, **kwargs, time_dependent=True)
        else:
            u_hat = u_hat.reshape((n * nt, modes[0]))
            u = basis(
                basis.transform(
                    basis.inv_transform(u_hat).reshape(
                        (n, nt, modes[0])
                    )
                ),
                **kwargs,
            )
        u.coeff = u.coeff.cpu()
        fst.coeff = fst.coeff.cpu()
        return (u, fst)
...
  \end{lstlisting}
  \caption{Implementation of Burgers generate function method of lines}\label{fig:burgers_generate_mol_impl}
\end{figure}

@@

        residual = u_t + uu_x - nu_u_xx - f
        return residual
...
  \end{lstlisting}
  \caption{Implementation of Burgers spectral residual function}\label{fig:burgers_spectral_residual_impl}
\end{figure}

\subsection{Testing}
\noindent The implementation is followed by tests of the functionality core to this study. The testing is done using the pytest library. The testing method used is a black box approach. The tests are written in a separate \verb|tests| directory. Each module is tested by a separate file. The tests are done to ensure compliance of the implementation to the requirements of the software. The results are presented in \lccref{table:impl_test_result}.

\begin{table}[H]
  \centering
  \begin{tabular}{lm{0.2\linewidth}m{0.5\linewidth}l}
    % \toprule
    \toprule
    No & Process & Test Target & Test Result       \\
    \midrule
    1  & Transform & Transform invertible with inverse transform & Achieved     \\\addlinespace[0.5em]
    2  & Evaluation & Transform coefficient evaluation close to original values & Achieved \\\addlinespace[0.5em]
    3  & Plot & Plotting function is run successfully & Achieved  \\\addlinespace[0.5em]
    4  & Perturb & Perturb function is run successfully & Achieved     \\\addlinespace[0.5em]
    5  & Operations & Operation functions is run successfully & Achieved     \\\addlinespace[0.5em]
    6  & Complex Coefficients Conversion & Conversion between real and complex coefficient representation is invertible for odd and even number of modes & Achieved     \\\addlinespace[0.5em]
    7  & SpectralSVR Train & Training function is run successfully & Achieved     \\\addlinespace[0.5em]
    8  & SpectralSVR Prediction & Coefficient prediction error within tolerable range & Achieved     \\\addlinespace[0.5em]
    9  & SpectralSVR Prediction & Function value prediction error within tolerable range & Achieved     \\
    \bottomrule
  \end{tabular}
  \caption{Testing results of the SpectralSVR Library implementation using black box method} \label{table:impl_test_result}
\end{table}

@@

\subsection{Scenario 1}
\noindent The first scenario

\subsection{Scenario 2}
\noindent

>>>>

\begin{figure}[H]
  \centering
  \begin{adjustbox}{max width=\textwidth}
    \tikzfig{figures/computational_model}
  \end{adjustbox}
  \caption{Computational Model of SpectralSVR.}\label{fig:computational_model}
\end{figure}

@@

[build-system]
requires = ["poetry-core"]
build-backend = "poetry.core.masonry.api"
  \end{lstlisting}
  \caption{Contents of pyproject.toml configuration file that define the dependencies using Poetry.}\label{fig:poetry_add_deps}
\end{figure}

@@

def to_complex_coeff(x: torch.Tensor) -> torch.Tensor:
  return torch.view_as_complex(
    x.reshape((*x.shape[:-1], -1, 2))
  )
  \end{lstlisting}
  \caption{Utility function to convert between complex matrices and the real representations.}\label{fig:complex_matrix_conversion_implementation}
\end{figure}

@@

    x_resized = torch.concat(
        (
            x_resized.index_select(dim, start_range),
            x_resized.index_select(dim, end_range),
        ),
        dim,
    )
  \end{lstlisting}
  \caption{Example of shrinking a tensor of samples' coefficients to a target number of modes.}\label{fig:coefficient_shrink_impl}
\end{figure}
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
for dim, (target_mode, current_mode) in enumerate(
    zip(target_modes, current_modes), 1
):
    device = x.device
    start_range = torch.tensor(
        range((current_mode - 1) // 2 + 1), dtype=torch.int
    ).to(device=device)
    # make sure that end range is empty if
    # the coefficient is only size 1
    end_range = torch.tensor(
        range(current_mode // 2, current_mode)
        if current_mode > 1
        else range(0),
        dtype=torch.int,
    ).to(device=device)
    padding_size = target_mode - current_mode
    modes = list(x_resized.shape)
    modes[dim] = padding_size
    padding = torch.zeros(modes).to(x_resized)

    x_resized = torch.concat(
        (
            x_resized.index_select(dim, start_range),
            padding,
            x_resized.index_select(dim, end_range),
        ),
        dim,
    )
  \end{lstlisting}
  \caption{Example of expanding a tensor of samples' coefficients to a target number of modes.}\label{fig:coefficient_expand_impl}
\end{figure}
Next, the second way to manage the size of a tensor along a dimension is to interpolate the current values. The implementation uses simple linear interpolation. This is especially useful for situations such as coefficients that are themselves functions of time. This is a common application such as when solving the initial condition problem of a function in space and time. One might represent the function as a Fourier series with basis as functions of space and coefficients as functions of time. In this case, the coefficients for time values that are not available can be interpolated. The implementation of this is shown in \lccref{fig:coefficient_interpolation_impl}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
    index_floor = index_float.floor().to(torch.int)
    index_ceil = index_float.ceil().to(torch.int)
    x_ceil = x.index_select(dim, index_ceil)
    x_floor = x.index_select(dim, index_floor)
    # interpolate coefficients
    index_shape = [1 for _ in range(x_floor.ndim)]
    index_shape[1] = -1
    index_scaler = (
        ((index_float - index_floor) / (
          index_ceil - index_floor
        ))
        .reshape(index_shape)
        .nan_to_num()
    )
    # ynt + scaler * (ynt1 - ynt)
    # (1 - scaler) * ynt + scaler * ynt1
    x_interp = torch.lerp(x_floor, x_ceil, index_scaler.to(x))
  \end{lstlisting}
  \caption{Example of interpolating for indices (floating point indices) between the available ones (intiger indices).}\label{fig:coefficient_interpolation_impl}
\end{figure}

@@

lsoda_solver: SolverSignatureType = partial(
    odeint, method="scipy_solver", options={"solver": "LSODA"}
)  # type: ignore
  \end{lstlisting}
  \caption{Implementation of ODE solvers and the types.}\label{fig:solver_impl}
\end{figure}

@@

        return F
...
  \end{lstlisting}
  \caption{Implementation of forward Fourier transform.}\label{fig:fourier_transform_impl}
\end{figure}

@@

        return F
...
  \end{lstlisting}
  \caption{Implementation of n-dimensional Fourier transform for a specific dimension.}\label{fig:ndim_fourier_transform_impl}
\end{figure}

@@

            F = torch.mm(f, e.T)
        return F
...
  \end{lstlisting}
  \caption{Implementation of the one dimensional Fourier Transform.}\label{fig:1d_fourier_transform_impl}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def __add__(self, other: Self):
        if isinstance(other, self.__class__):
            if other.coeff is None:
                return self.copy()
            elif self.coeff is None:
                return other.copy()
            else:
                result = self.resize_modes(other)
                result.coeff = result.coeff + other.coeff
                return result
        else:
            raise TypeError(
                f"unsupported operand type(s) for +: '{
                    self.__class__
                }' and '{type(other)}'"
            )
...
  \end{lstlisting}
  \caption{Implementation of basis addition function.}\label{fig:basis_addition_impl}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def grad(self, dim: int = 0, ord: int = 1) -> Self:
        copy = self.copy()
        if dim == 0 and self.time_dependent:
            # time dependent use finite differences
            dt = self.periods[0] / (self.time_size - 1)
            coeff = copy.coeff
            for o in range(ord):
                coeff = torch.gradient(coeff, spacing=dt, dim=1)[0]
            copy.coeff = coeff
        else:
            if self.time_dependent:
                # disregard time dimension
                dim = dim - 1
            k = copy.wave_number(copy.modes[dim])
            multiplier_dims = [1 for _ in range(copy.ndim)]
            multiplier_dims[dim] = copy.modes[dim]
            if self.time_dependent:
                multiplier_dims = (1, *multiplier_dims)
            multiplier = (
                2
                * torch.pi
                * 1j
                * k.reshape(multiplier_dims).to(copy.coeff)
                / self.periods[dim]
            )
            multiplier = multiplier.pow(ord)
            coeff = copy.coeff.mul(multiplier)
            coeff[:, ..., 0] = torch.tensor(0 + 0j)
            copy.coeff = coeff
        return copy
...
  \end{lstlisting}
  \caption{Implementation of basis grad function.}\label{fig:basis_grad_impl}
\end{figure}

@@

            vals = cls.inv_transform(coeff)
            coeff = cls.transform(vals.real + 0j)
        if scale:
            scaler = torch.tensor(modes).sum() * 0.2
            coeff = coeff.mul(scaler)
        return coeff
...
  \end{lstlisting}
  \caption{Implementation of Fourier Basis coefficient generation function.}\label{fig:generate_coeff_impl}
\end{figure}

@@

\subsubsection{Model}\label{sec:impl_model}

@@

        return (b, alpha)
...
  \end{lstlisting}
  \caption{Implementation of Least-squares Support Vector Regression fitting function.}\label{fig:lssvr_fit_function_impl}
\end{figure}

@@

        if X.ndim == 1:
            return predictions.reshape(-1)
        else:
            return predictions
...
  \end{lstlisting}
  \caption{Implementation of Least-squares Support Vector Regression prediction function.}\label{fig:lssvr_predict_function_impl}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def get_correlation_image(self):
        return self._batched_K(
            self.sv_x, self.sv_x
        ).mm(self.sv_x)
...
  \end{lstlisting}
  \caption{Implementation of Least-squares Support Vector Regression correlation image function.}\label{fig:lssvr_correlation_image_function_impl}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def get_p_matrix(self):
        return self.sv_x.T.mm(self.alpha)
...
  \end{lstlisting}
  \caption{Implementation of Least-squares Support Vector Regression p-matrix function.}\label{fig:lssvr_p_matrix_function_impl}
\end{figure}

@@

        if torch.is_complex(u_coeff):
            u_coeff = to_real_coeff(u_coeff)
        if torch.is_complex(f):
            f = to_real_coeff(f)
        self.svr.fit(f, u_coeff)
        return self
...
  \end{lstlisting}
  \caption{Implementation of SpectralSVR training function.}\label{fig:spectralsvr_train_impl}
\end{figure}

@@

        return self.basis.evaluate(
            coeff=coeff.reshape((f.shape[0], *self.modes)),
            x=x,
            periods=periods,
            time_dependent=self.basis.time_dependent,
        )
...
  \end{lstlisting}
  \caption{Implementation of SpectralSVR forward function (pointwise prediction).}\label{fig:spectralsvr_forward_impl}
\end{figure}

The final core piece of functionality the SpectralSVR class provides is the inverse parameter estimation. This is implemented to solve inverse problems such as predicting the initial conditions for the Burgers' equation or heat equation. This function essentially does the gradient descent on a loss function in such a way that the output function matches some expected output. The loss function is simply a measure of how far the outputs of the current predicted inputs are from the target outputs. The loss function is formulated as the mean squared error of the difference between the predicted output and expected output. Using some randomly initialized values as the inputs, the predicted output is then used to compute the loss. The gradient of the loss with respect to the inputs are then computed and then used to perform optimization of the inputs using the ADAM optimizer. This is implemented as the inverse function shown in \lccref{fig:spectralsvr_inverse_impl}. The function arguments are the expected output function coefficients, evaluation points, the loss function used, how many optimization loops are done, the random number generator, and the gain which is used in generating the random initial input functions. The function returns the estimated input function coefficients in the real representation. For Fourier coefficients this means that the predicted input coefficients need to first be converted to complex valued tensors and then used to construct a FourierBasis instance. For convenience, a wrapper like the forward function is also implemented for evaluating the function values of the predicted input coefficients.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def inverse_coeff(
        self,
        u_coeff: torch.Tensor,
        loss_fn: Callable[
            [torch.Tensor, torch.Tensor], torch.Tensor
        ] = mean_squared_error,
        epochs=100,
        generator=torch.Generator().manual_seed(42),
        gain=0.05,
        **optimizer_params,
    ):
        f_shape = (u_coeff.shape[0], self.svr.sv_x.shape[1])
        complex_coeff = u_coeff.is_complex()
        original_device = u_coeff.device
        u_coeff = to_real_coeff(u_coeff.flatten(1)).to(self.svr.device)

        # inverse problem
        f_coeff_pred = (
            torch.randn(f_shape, generator=generator).to(self.svr.device) * gain
        )
        f_coeff_pred.requires_grad_()
        optim = torch.optim.Adam([f_coeff_pred], **optimizer_params)

        for epoch in range(epochs):
            optim.zero_grad()
            u_coeff_pred = self.svr.predict(f_coeff_pred)
            loss = loss_fn(u_coeff_pred, u_coeff)
            loss.backward()
            optim.step()
        optim.zero_grad()
        f_coeff_pred.requires_grad_(False)
        if complex_coeff:
            f_coeff_pred = to_complex_coeff(f_coeff_pred)
        return f_coeff_pred.to(original_device)
...
  \end{lstlisting}
  \caption{Implementation of SpectralSVR inverse coeff function (parameter estimation).}\label{fig:spectralsvr_inverse_impl}
\end{figure}

The SpectralSVR class also provides a convenience function for testing the performance of the learned model. This function is shown in \lccref{fig:spectralsvr_test_impl}. The function arguments are the testing inputs, expected testing outputs, and a resolution for evaluation of function values. The first step is to convert the testing input into real representations and then predicting the output function. The predicted output are then compared to the testing outputs using many metrics including RMSE, MSE, MAE, R2, SMAPE, RSE, RRSE, and the number of Nan values present in the predicted output. The same comparison process is carried out again for the function values which are evaluated at the specified resolution. Finally, the function returns the metric values as a nested tuple for both the coefficients predictions and the function value predictions.

@@

        grid = self.basis.grid(res).flatten(0, -2)
        u_preds = self.basis.evaluate(
            coeff=to_complex_coeff(u_coeff_preds),
            x=grid,
            time_dependent=self.basis.time_dependent,
        ).real
        u_targets = self.basis.evaluate(
            coeff=to_complex_coeff(u_coeff_targets),
            x=grid,
            time_dependent=self.basis.time_dependent,
        ).real
        return {
            "spectral": get_metrics(
                u_coeff_preds, u_coeff_targets
            ),
            "function value": get_metrics(u_preds, u_targets),
        }
...
  \end{lstlisting}
  \caption{Implementation of SpectralSVR test function.}\label{fig:spectralsvr_test_impl}
\end{figure}

@@

    @abc.abstractmethod
    def residual(
        self, u: BasisSubType, *args, **kwargs
    ) -> BasisSubType:
        pass
  \end{lstlisting}
  \caption{Implementation of Problem base class.}\label{fig:problem_impl}
\end{figure}

@@

        return (u, ut)
...
  \end{lstlisting}
  \caption{Implementation of Antiderivative generate function.}\label{fig:antiderivative_generate_impl}
\end{figure}

@@

    def residual(
        self, u: BasisSubType, ut: BasisSubType
    ) -> BasisSubType:
        u_val, grid = u.get_values_and_grid()
        ut_val = ut.get_values()
        dt = grid[1, 0] - grid[0, 0]
        u_grad = torch.gradient(
            u_val, spacing=dt.item(), dim=1
        )[0]
        residual_val = u_grad - ut_val
        residual = u.copy()
        residual.coeff = u.transform(residual_val)
        return residual
...
  \end{lstlisting}
  \caption{Implementation of Antiderivative residual functions.}\label{fig:antiderivative_residual_impl}
\end{figure}

@@

        L = space_domain.stop - space_domain.start
        x = (
            basis.grid(slice(
                space_domain.start, space_domain.stop, modes[0]
            )).flatten(0, -2)
            .to(device=device)
        )
        T: float = time_domain.stop - time_domain.start
        nt = int(time_domain.step)
        # nt = int(T / (0.01 * nu) + 2)
        dt = T / (nt - 1)
        t = basis.grid(time_domain).flatten().to(device=device)
        periods = (T, L)
...
  \end{lstlisting}
  \caption{Implementation of Burgers generate function.}\label{fig:burgers_generate_impl}
\end{figure}

@@

                f_val = fst.get_values(
                    res=(time_domain, *res_modes)
                )
                f_coeff = basis.transform(
                    f_val.flatten(0, 1)
                ).reshape((n, nt, *modes))
                f_gen = basis(
                    coeff=f_coeff,
                    time_dependent=True,
                    periods=periods
                )
...
  \end{lstlisting}
  \caption{Implementation of Burgers generate function manufactured method.}\label{fig:burgers_generate_manufactured_impl}
\end{figure}

@@

        u_u_x_hat = uu_x.coeff
        u_xx_hat = u.grad().grad().coeff
        u_t_hat = nu * u_xx_hat + f_hat - u_u_x_hat
        return u_t_hat
...
  \end{lstlisting}
  \caption{Implementation of Burgers generate function time derivative.}\label{fig:burgers_generate_dt_impl}
\end{figure}
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
        u0.coeff = u0.coeff.to(device=device)
        fst.coeff = fst.coeff.to(device=device)
        print(f"generating with {len(t)} time steps")

@@

        u_hat = u_hat.movedim(0, 1)
        if timedependent_solution:
            u = basis(u_hat, **kwargs, time_dependent=True)
        else:
            u_hat = u_hat.reshape((n * nt, modes[0]))
            u = basis(
                basis.transform(
                    basis.inv_transform(u_hat).reshape(
                        (n, nt, modes[0])
                    )
                ),
                **kwargs,
            )
        u.coeff = u.coeff.cpu()
        fst.coeff = fst.coeff.cpu()
        return (u, fst)
...
  \end{lstlisting}
  \caption{Implementation of Burgers generate function method of lines.}\label{fig:burgers_generate_mol_impl}
\end{figure}

@@

        residual = u_t + uu_x - nu_u_xx - f
        return residual
...
  \end{lstlisting}
  \caption{Implementation of Burgers spectral residual function.}\label{fig:burgers_spectral_residual_impl}
\end{figure}

\subsection{Testing}
\noindent The implementation is followed by tests of the functionality core to this study. The testing is done using the pytest library. The testing method used is a black box approach. The tests are written in a separate \verb|tests| directory. Each module is tested by a separate file. The tests are done to ensure compliance of the implementation to the requirements of the software. The final results are presented in \lccref{table:impl_test_result}. Each version and iteration of the development must ensure that the tests are passed before any changes are committed to the Github repository.

\begin{table}[H]
  \centering
  \begin{tabular}{lm{0.2\linewidth}m{0.5\linewidth}l}
    % \toprule
    \toprule
    No & Process & Test Target & Test Result       \\
    \midrule
    1  & Transform & Transform invertible with inverse transform & Achieved     \\\addlinespace[0.5em]
    2  & Evaluation & Transform coefficient evaluation close to original values & Achieved \\\addlinespace[0.5em]
    3  & Plot & Plotting function is run successfully & Achieved  \\\addlinespace[0.5em]
    4  & Perturb & Perturb function is run successfully & Achieved     \\\addlinespace[0.5em]
    5  & Operations & Operation functions is run successfully & Achieved     \\\addlinespace[0.5em]
    6  & Complex Coefficients Conversion & Conversion between real and complex coefficient representation is invertible for odd and even number of modes & Achieved     \\\addlinespace[0.5em]
    7  & SpectralSVR Train & Training function is run successfully & Achieved     \\\addlinespace[0.5em]
    8  & SpectralSVR Prediction & Coefficient prediction error within tolerable range & Achieved     \\\addlinespace[0.5em]
    9  & SpectralSVR Prediction & Function value prediction error within tolerable range & Achieved     \\
    \bottomrule
  \end{tabular}
  \caption{Testing results of the SpectralSVR Library implementation using black box method.} \label{table:impl_test_result}
\end{table}

@@

\subsection{Scenario 1}
\noindent The first scenario results are presented in four parts. The first is the coefficient prediction metrics which is shown in \lccref{table:scenario_1_spectral_metrics}. The table shows the noise level, the corresponding kernel scaling factor sigma, and the performance metrics themselves.
\begin{table}[H]
  \caption{Performance metrics of coefficient prediction in scenario 1 by noise level.}\label{table:scenario_1_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrrrrrr}
    \toprule
    Noise level & Sigma & MSE & RMSE & MAE & R2 & sMAPE \\
    \midrule
    5\% & 10.00 & 0.85 & 0.92 & 0.56 & 0.96 & 0.31 \\
    10\% & 10.00 & 2.07 & 1.44 & 0.93 & 0.91 & 0.46 \\
    50\% & 10.00 & 15.56 & 3.94 & 3.04 & 0.50 & 1.04 \\
    \bottomrule
  \end{tabular}
\end{table}
